---
title: "Modeling"
author: "Eliza Baier"
date: "March 8, 2025"
format: 
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    toc-float: true
    toc-title: "Contents"
    self-contained: true
execute:
  include: true
  eval: true
  warning: false
  message: false
editor: 
  markdown: 
    wrap: sentence
---

## Business Problem Summary

Unbanked individuals struggle to get loans from trustworthy lenders which prevents them from owning property and establishing credit history. To properly serve this population, Home Credit needs to be able to predict their potential clients’ risk of defaulting. Accurately predicting the risk associated with each client will allow Home Credit to effectively serve an underserved population while increasing revenue and minimizing risk. 

## Analytical Problem Summary

The analytical challenge with this project is to take a highly-dimensional dataset and create a model that is not biased against the target underserved populations. To do this, it will be important to use a combination of the provided credit scores and transactional data for those who do not have a credit score. To decrease the dimensionality of the dataset, it will be important to eliminate some of the columns and clean the data so that the remaining columns are an accurate representation of the sample.

## Modeling Approach to the Business Problem

This notebook contains my individual modeling efforts the provided data. This modeling aims to build a baseline model as well as several Random Forest models that I hope will outperform the baseline. To do this, I first loaded and examined the application data, combined it with the bureau data, cleaned and evaluated the merged dataset, split the training data into train and validation folds to provide an avenue for evaluating and comparing model metrics, and built both a baseline model and six random forest models.

## Load Libraries

First, I loaded necessary libraries and read in the application data csv. 

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(randomForest)
library(caret)
library(nnet)
library(rminer)
train <- read.csv("application_train.csv")
```

## Clean and Update Application Data

The first step after loading libraries was to clean and update the application data. I used feature engineering to create several new variables and removed all of the irrelevant variables that I identified in my EDA from the dataset. I then cleaned and imputed important variables with appropriate values and finished preparing the dataset to be merged with the bureau data. 

```{r}
# Copy train set to new variable
train_clean <- train

# Impute OWN_CAR_BIN NAs with -1
train_clean <- train_clean |>
  mutate(OWN_CAR_AGE = ifelse(is.na(OWN_CAR_AGE), -1, OWN_CAR_AGE))

# Create categorical OWN_CAR_BIN
train_clean$OWN_CAR_BIN <- train_clean$OWN_CAR_AGE |>
        cut(breaks = c(-1, 0, 5, 10, 15, 20, 25, 30, 100), 
        right = FALSE, 
        labels = c("No Car", "0-4 Years Old", "5-9 Years Old", "10-14 Years Old", 
                 "15-19 Years Old", "20-24 Years Old", "25-29 Years Old", "30+ Years Old"))

# Check OWN_CAR_BIN variable
train_clean |>
  count(OWN_CAR_BIN) |>
  mutate(percentage = n / sum(n) * 100) |>
  arrange(desc(n))

# Create binary OWN_ASSET variable
train_clean <- train_clean |>
  mutate(OWN_ASSET = ifelse(FLAG_OWN_REALTY == "Y" | FLAG_OWN_CAR == "Y", 1, 0))

# Check OWN_ASSET variable
train_clean |>
  count(OWN_ASSET) |>
  mutate(percentage = n / sum(n) * 100) |>
  arrange(desc(n))

# Remove irrelevant variables
train_clean <- train_clean |>
  select(-c(COMMONAREA_AVG, COMMONAREA_MODE, COMMONAREA_MEDI, NONLIVINGAPARTMENTS_AVG, NONLIVINGAPARTMENTS_MODE, NONLIVINGAPARTMENTS_MEDI, LIVINGAPARTMENTS_AVG, LIVINGAPARTMENTS_MODE, LIVINGAPARTMENTS_MEDI, FLOORSMIN_AVG, FLOORSMIN_MODE, FLOORSMIN_MEDI, YEARS_BUILD_AVG, YEARS_BUILD_MODE, YEARS_BUILD_MEDI, LANDAREA_AVG, LANDAREA_MODE, LANDAREA_MEDI, BASEMENTAREA_AVG, BASEMENTAREA_MODE, BASEMENTAREA_MEDI, NONLIVINGAREA_AVG, NONLIVINGAREA_MODE, NONLIVINGAREA_MEDI, ELEVATORS_AVG, ELEVATORS_MODE, ELEVATORS_MEDI, APARTMENTS_AVG, APARTMENTS_MODE, APARTMENTS_MEDI, ENTRANCES_AVG, ENTRANCES_MODE, ENTRANCES_MEDI, LIVINGAREA_AVG, LIVINGAREA_MODE, LIVINGAREA_MEDI, CNT_CHILDREN, FLAG_MOBIL, FLAG_EMP_PHONE, FLAG_WORK_PHONE, FLAG_CONT_MOBILE, FLAG_EMAIL, REG_REGION_NOT_LIVE_REGION, REG_REGION_NOT_WORK_REGION, LIVE_REGION_NOT_WORK_REGION, REG_CITY_NOT_LIVE_CITY, LIVE_CITY_NOT_WORK_CITY, YEARS_BEGINEXPLUATATION_AVG, YEARS_BEGINEXPLUATATION_MODE, YEARS_BEGINEXPLUATATION_MEDI, EMERGENCYSTATE_MODE, DEF_30_CNT_SOCIAL_CIRCLE, OBS_30_CNT_SOCIAL_CIRCLE, OBS_60_CNT_SOCIAL_CIRCLE, DEF_60_CNT_SOCIAL_CIRCLE, FLAG_DOCUMENT_2, FLAG_DOCUMENT_3, FLAG_DOCUMENT_4, FLAG_DOCUMENT_5, FLAG_DOCUMENT_6, FLAG_DOCUMENT_7, FLAG_DOCUMENT_8, FLAG_DOCUMENT_9, FLAG_DOCUMENT_10, FLAG_DOCUMENT_11, FLAG_DOCUMENT_12, FLAG_DOCUMENT_13, FLAG_DOCUMENT_14, FLAG_DOCUMENT_15, FLAG_DOCUMENT_16, FLAG_DOCUMENT_17, FLAG_DOCUMENT_18, FLAG_DOCUMENT_19, FLAG_DOCUMENT_20, FLAG_DOCUMENT_21, AMT_REQ_CREDIT_BUREAU_HOUR, AMT_REQ_CREDIT_BUREAU_DAY, AMT_REQ_CREDIT_BUREAU_WEEK, AMT_REQ_CREDIT_BUREAU_MON, AMT_REQ_CREDIT_BUREAU_QRT, AMT_REQ_CREDIT_BUREAU_YEAR, FLAG_PHONE, FLOORSMAX_AVG, FLOORSMAX_MODE, FLOORSMAX_MEDI, HOUR_APPR_PROCESS_START, REG_CITY_NOT_WORK_CITY, TOTALAREA_MODE, WALLSMATERIAL_MODE, FONDKAPREMONT_MODE, WEEKDAY_APPR_PROCESS_START, REGION_RATING_CLIENT_W_CITY))

# Factor remaining categorical variables
train_clean <- train_clean |>
  mutate(NAME_CONTRACT_TYPE = factor(NAME_CONTRACT_TYPE),
         NAME_TYPE_SUITE = factor(NAME_TYPE_SUITE),
         OCCUPATION_TYPE = factor(OCCUPATION_TYPE),
         CODE_GENDER = factor(CODE_GENDER),
         FLAG_OWN_CAR = factor(FLAG_OWN_CAR),
         FLAG_OWN_REALTY = factor(FLAG_OWN_REALTY),
         NAME_TYPE_SUITE = factor(NAME_TYPE_SUITE),
         NAME_INCOME_TYPE = factor(NAME_INCOME_TYPE),
         NAME_EDUCATION_TYPE = factor(NAME_EDUCATION_TYPE),
         NAME_FAMILY_STATUS = factor(NAME_FAMILY_STATUS),
         NAME_HOUSING_TYPE = factor(NAME_HOUSING_TYPE),
         HOUSETYPE_MODE = factor(HOUSETYPE_MODE),
         OCCUPATION_TYPE = factor(OCCUPATION_TYPE),
         ORGANIZATION_TYPE = factor(ORGANIZATION_TYPE),
         HOUSETYPE_MODE = factor(HOUSETYPE_MODE),
         OWN_ASSET = factor(OWN_ASSET),
         TARGET = factor(TARGET))

# Combine 'Other' categories in factored variables 
train_clean <- train_clean |>
  mutate(
    NAME_TYPE_SUITE = case_when(
      NAME_TYPE_SUITE %in% c("", "Other_A", "Other_B", "Other") ~ "Other",
      # Combine blanks and 2 "other" categories into "Other" category
      TRUE ~ NAME_TYPE_SUITE
    ),
    # Keep all other values unchanged
    OCCUPATION_TYPE = case_when(OCCUPATION_TYPE == "" ~ "Other", # Replace blank values with "Other"
                                TRUE ~ OCCUPATION_TYPE),
    # Keep all other values unchanged
    HOUSETYPE_MODE = case_when(HOUSETYPE_MODE == "" ~ "other", # Replace blank values with "Other"
                               TRUE ~ HOUSETYPE_MODE))

# Impute DAYS_EMPLOYED with median
train_clean <- train_clean |>
  mutate(DAYS_EMPLOYED = ifelse(DAYS_EMPLOYED > 18250, median(DAYS_EMPLOYED), DAYS_EMPLOYED))

# Impute EXT_SOURCE_1 with 0 indicating no credit score
train_clean <- train_clean |>
  mutate(EXT_SOURCE_1 = ifelse(is.na(EXT_SOURCE_1), 0, EXT_SOURCE_1))

# Impute EXT_SOURCE_2 with 0 indicating no credit score
train_clean <- train_clean |>
  mutate(EXT_SOURCE_2 = ifelse(is.na(EXT_SOURCE_2), 0, EXT_SOURCE_2))

# Impute EXT_SOURCE_3 with 0 indicating no credit score
train_clean <- train_clean |>
  mutate(EXT_SOURCE_3 = ifelse(is.na(EXT_SOURCE_3), 0, EXT_SOURCE_3))

# Impute CNT_FAM_MEMBERS with the median
train_clean <- train_clean |>
  mutate(CNT_FAM_MEMBERS = ifelse(is.na(CNT_FAM_MEMBERS), median(CNT_FAM_MEMBERS, na.rm = TRUE), CNT_FAM_MEMBERS))

# Impute AMT_ANNUITY with the median
train_clean <- train_clean |>
  mutate(AMT_ANNUITY = ifelse(is.na(AMT_ANNUITY), median(AMT_ANNUITY, na.rm = TRUE), AMT_ANNUITY))

# Impute AMT_GOODS_PRICE with the median
train_clean <- train_clean |>
  mutate(AMT_GOODS_PRICE = ifelse(is.na(AMT_GOODS_PRICE), median(AMT_GOODS_PRICE, na.rm = TRUE), AMT_GOODS_PRICE))

# Impute DAYS_LAST_PHONE_CHANGE with the median
train_clean <- train_clean |>
  mutate(DAYS_LAST_PHONE_CHANGE = ifelse(is.na(DAYS_LAST_PHONE_CHANGE), median(DAYS_LAST_PHONE_CHANGE, na.rm = TRUE), DAYS_LAST_PHONE_CHANGE))

colSums(is.na(train_clean))

train_clean |>
  str()

# Save as RDS
saveRDS(train_clean, "application_train_cleaned.rds")

# Save as CSV for visual exploration
write.csv(train_clean, "application_train_cleaned.csv", row.names = FALSE)
```

## Read in Cleaned Application and Bureau Data

Another team member cleaned the bureau data and I read it in to my notebook using the readRDS function.

```{r}
getwd()
setwd("C:\\Users\\eliza\\Documents\\MSBA\\Spring 2025\\IS 6812 Practice Capstone")
file.exists("application_train_cleaned.csv")
application <- train_clean
bureau <- readRDS("mutated_merged_bureau.rds")
head(bureau)
```

## Merge Bureau Data with Application Data

I then merged the cleaned bureau data with my cleaned application data and checked for missings. Because the bureau dataset had a lot of missing values, I was able to identify columns that would need to be imputed. 

```{r}
# Merge application and bureau data
merge <- application |>
  left_join(bureau, by = "SK_ID_CURR")

# Check merge
head(merge)

# Get number of missings
sum(is.na(merge$SK_ID_BUREAU))

# Summarize missings
na_summary <- merge |> 
  summarise(across(everything(), ~ sum(is.na(.)))) |>
  pivot_longer(everything(), names_to = "column", values_to = "missing_count") |>
  filter(missing_count > 0) 

# Get column list for missing data
na_list <- setNames(as.list(na_summary$missing_count), na_summary$column) 
na_list
```

```{r}
# Check join
length(intersect(application$SK_ID_CURR, bureau$SK_ID_CURR))
length(unique(application$SK_ID_CURR))
length(unique(bureau$SK_ID_CURR))
```

## Impute the Bureau NAs with 0

The bureau data contained many missing values so I imputed the vast majority of them with 0s, and a few others with appropriate values. 

```{r}
# Fill in missing data with 0s
impute_numeric_col <- function(df, col_name, strategy = "median") {
  # Check if the specified column exists
  if (!col_name %in% names(df)) {
    warning(paste("Column", col_name, "not found in df. Returning original df."))
    return(df)
  }
  # Check if it's numeric
  if (!is.numeric(df[[col_name]])) {
    warning(paste("Column", col_name, "is not numeric. Returning original df."))
    return(df)
  }
  
  # Determine the fill value based on the strategy
  fill_val <- NA
  if (strategy == "zero") {
    fill_val <- 0
  } else if (strategy == "min") {
    fill_val <- min(df[[col_name]], na.rm = TRUE)
  } else if (strategy == "max") {
    fill_val <- max(df[[col_name]], na.rm = TRUE)
  } else if (strategy == "median") {
    fill_val <- median(df[[col_name]], na.rm = TRUE)
  } else if (strategy == "mean") {
    fill_val <- mean(df[[col_name]], na.rm = TRUE)
  } else if (strategy == "sum") {
    fill_val <- sum(df[[col_name]], na.rm = TRUE)
  } else {
    stop("strategy must be one of 'zero', 'min', 'max', 'median', or 'mean'")
  }
  
  # Replace NA with the chosen fill value
  df[[col_name]][is.na(df[[col_name]])] <- fill_val
  
  return(df)
}

# Impute NAs in merge$SK_ID_BUREAU with 0s
merge <- merge |>
  mutate(SK_ID_BUREAU = ifelse(is.na(SK_ID_BUREAU), 0, SK_ID_BUREAU))

# Remove AMT_ANNUITY.y and CREDIT_CURRENCY because of lack of contribution to the model
merge <- merge |>
  select(-AMT_ANNUITY.y, -CREDIT_CURRENCY)

# Impute NAs in CREDIT_TYPE with "None"
merge <- merge |>
  mutate(CREDIT_TYPE = ifelse(is.na(CREDIT_TYPE), "None", CREDIT_TYPE))

# Impute NAs in STATUS with "X"
merge <- merge |>
  mutate(STATUS = ifelse(is.na(STATUS), "X", STATUS))

# Impute NAs in DAYS_CREDIT with 0
merge <- merge |>
  mutate(DAYS_CREDIT = ifelse(is.na(DAYS_CREDIT), 0, DAYS_CREDIT))

# Impute NAs in CREDIT_DAY_OVERDUE with 0
merge <- merge |>
  mutate(CREDIT_DAY_OVERDUE = ifelse(is.na(CREDIT_DAY_OVERDUE), 0, CREDIT_DAY_OVERDUE))

# Impute NAs in DAYS_CREDIT_ENDDATE with 0
merge <- merge |>
  mutate(DAYS_CREDIT_ENDDATE = ifelse(is.na(DAYS_CREDIT_ENDDATE), 0, DAYS_CREDIT_ENDDATE))

# Impute NAs in DAYS_ENDDATE_FACT with 0
merge <- merge |>
  mutate(DAYS_ENDDATE_FACT = ifelse(is.na(DAYS_ENDDATE_FACT), 0, DAYS_ENDDATE_FACT))

# Impute NAs in AMT_CREDIT_MAX_OVERDUE with 0
merge <- merge |>
  mutate(AMT_CREDIT_MAX_OVERDUE = ifelse(is.na(AMT_CREDIT_MAX_OVERDUE), 0, AMT_CREDIT_MAX_OVERDUE))

# Impute NAs in CNT_CREDIT_PROLONG with 0
merge <- merge |>
  mutate(CNT_CREDIT_PROLONG = ifelse(is.na(CNT_CREDIT_PROLONG), 0, CNT_CREDIT_PROLONG))

# Impute NAs in AMT_CREDIT_SUM with 0
merge <- merge |>
  mutate(AMT_CREDIT_SUM = ifelse(is.na(AMT_CREDIT_SUM), 0, AMT_CREDIT_SUM))

# Impute NAs in AMT_CREDIT_SUM_DEBT with 0
merge <- merge |>
  mutate(AMT_CREDIT_SUM_DEBT = ifelse(is.na(AMT_CREDIT_SUM_DEBT), 0, AMT_CREDIT_SUM_DEBT))

# Impute NAs in AMT_CREDIT_SUM_LIMIT with 0
merge <- merge |>
  mutate(AMT_CREDIT_SUM_LIMIT = ifelse(is.na(AMT_CREDIT_SUM_LIMIT), 0, AMT_CREDIT_SUM_LIMIT))

# Impute NAs in AMT_CREDIT_SUM_OVERDUE with 0
merge <- merge |>
  mutate(AMT_CREDIT_SUM_OVERDUE = ifelse(is.na(AMT_CREDIT_SUM_OVERDUE), 0, AMT_CREDIT_SUM_OVERDUE))

# Impute NAs in DAYS_CREDIT_UPDATE with 0
merge <- merge |>
  mutate(DAYS_CREDIT_UPDATE = ifelse(is.na(DAYS_CREDIT_UPDATE), 0, DAYS_CREDIT_UPDATE))

# Impute NAs in CREDIT_ACTIVE with 0
merge <- merge |>
  mutate(CREDIT_ACTIVE = ifelse(is.na(CREDIT_ACTIVE), 0, CREDIT_ACTIVE))

# Impute NAs in MONTHS_BALANCE with 0
merge <- merge |>
  mutate(MONTHS_BALANCE = ifelse(is.na(MONTHS_BALANCE), 0, MONTHS_BALANCE))
```

## Downsample

After loading, cleaning, and merging the data, I down-sampled on the target variable to create a more even target distribution and so that I could use a subset of the data which would help my my Random Forest models would run faster. 

```{r}
# Create downsample function
downsample_target <- function(df, target_col, ratio = 1.0, seed = 42) {
  # Convert target_col to string if it's a symbol
  target_col <- rlang::ensym(target_col)  # if you want tidy evaluation
  
  # Separate majority/minority
  minority_df <- df %>% dplyr::filter(!!target_col == 1)
  majority_df <- df %>% dplyr::filter(!!target_col == 0)
  
  set.seed(seed)
  # Desired majority size
  desired_majority_size <- floor(ratio * nrow(minority_df))
  
  # Downsample majority
  majority_downsampled <- majority_df %>%
    dplyr::sample_n(size = desired_majority_size)
  
  # Combine
  combined_df <- dplyr::bind_rows(minority_df, majority_downsampled)
  
  return(combined_df)
}

# Downsample merged dataset
merge_ds <- downsample_target(merge, TARGET, ratio = 2.33, seed = 123)

# Target variable distribution in newdownsampled dataset
merge_ds |>
  count(TARGET) |>
  mutate(percentage = n / sum(n) * 100) |>
  arrange(desc(n))
```

## Train and Evaluate Logistic Regression

I first trained and evaluated a logistic regression as a baseline to compare future models against. This model performed fairly well, with an AUC of 0.70. 

```{r}
# Fit logistic regression model
model1 <- glm(TARGET ~ NAME_CONTRACT_TYPE + OWN_CAR_BIN + DAYS_EMPLOYED + NAME_EDUCATION_TYPE + EXT_SOURCE_1 +  EXT_SOURCE_2 + EXT_SOURCE_3, data = merge_ds, family = binomial)

# View model summary
summary(model1)

# Predict probabilities on the same dataset
pred_probs <- predict(model1, newdata = test_data, type = "response")

# Compute the ROC curve
roc_curve <- roc(test_data$TARGET, pred_probs)

# Get the AUC score
auc_score <- auc(roc_curve)

# Print AUC for the logistic regression model
print(round(auc_score, 2))
```

## Train & Evaluate Random Forest Model

I trained six random forest models in the hopes that they would perform better than our benchmark logistic model. Unfortunately, because of the extreme imbalance in the TARGET variable, the random forest ended up chronically overfitting on the train set and therefore performing very poorly on the test set, despite downsampling to try and prevent this. Despite trying several different variations of the model, I was unable to solve the overfitting issue so my group moved on to trying gradient boosting.

```{r}
Train the Random Forest model 2
set.seed(123)
model2 <- randomForest(TARGET ~ . -ORGANIZATION_TYPE, data = train_merge_ds,
                   	ntree = 500,
                   	mtry = sqrt(ncol(train_merge_ds) - 1),
                   	importance = TRUE)

# Evaluate model 2
# View sorted importance of each variable
importance_values <- importance(model2)  # Get feature importance scores
importance_values
sorted_importance <- importance_values[order(-importance_values[, 1]), ]  # Sort by first column (MeanDecreaseGini)
sorted_importance

# Make predictions for model 2 on the train set
train_predictions_m2 <- predict(model2, newdata = train_merge_ds, type = "response")
train_predictions_m2 <- as.numeric(as.character(train_predictions_m2)) # convert to numeric to calculate ROC

# Make predictions for model 2 on the test set
test_predictions_m2 <- predict(model2, newdata = validation_merge_ds, type = "response")
test_predictions_m2 <- as.numeric(as.character(test_predictions_m2)) # convert to numeric to calculate ROC

# Calculate AUC for model 2 for train set
train_roc_m2 <- roc(train_merge_ds$TARGET, train_predictions_m2)
train_auc_m2 <- auc(train_roc_m2)
print(paste("AUC for train set:", train_auc_m2)) # AUC = 1.00

# Calculate AUC for model 2 for test set
test_roc_m2 <- roc(validation_merge_ds$TARGET, test_predictions_m2)
test_auc_m2 <- auc(test_roc_m2)
print(paste("AUC for test set:", test_auc_m2)) # AUC = 0.61


# Train Random Forest model 3
model3 <- randomForest(TARGET ~ . -ORGANIZATION_TYPE, data = train_merge_ds,
                   	ntree = 500,
                   	mtry = sqrt(ncol(train_merge_ds) - 1),
                   	importance = TRUE,
                   	classwt = c(0.43, 2.33))

# Evaluate model 3
# Make predictions for model 3 on the train set
train_predictions_m3 <- predict(model3, newdata = train_merge_ds, type = "response")
train_predictions_m3 <- as.numeric(as.character(train_predictions_m3)) # convert to numeric to calculate ROC

# Make predictions for model 3 on the test set
test_predictions_m3 <- predict(model3, newdata = validation_merge_ds, type = "response")
test_predictions_m3 <- as.numeric(as.character(test_predictions_m3)) # convert to numeric to calculate ROC

# Calculate AUC for model 3 for train set
train_roc_m3 <- roc(train_merge_ds$TARGET, train_predictions_m3)
train_auc_m3 <- auc(train_roc_m3)
print(paste("AUC for train set:", train_auc_m3)) # AUC = 1.00

# Calculate AUC for model 3 for test set
test_roc_m3 <- roc(validation_merge_ds$TARGET, test_predictions_m3)
test_auc_m3 <- auc(test_roc_m3)
print(paste("AUC for test set:", test_auc_m3)) # AUC = 0.58

# Train Random Forest model 4
model4 <- randomForest(TARGET ~ . -ORGANIZATION_TYPE, data = train_merge_ds,
                   	ntree = 200,
                   	maxnodes = 30,
                   	nodesize = 10,
                   	mtry = sqrt(ncol(train_merge_ds) - 1),
                   	importance = TRUE)
# Evaluate model 4
# Make predictions for model 4 on the train set
train_predictions_m4 <- predict(model4, newdata = train_merge_ds, type = "response")
train_predictions_m4 <- as.numeric(as.character(train_predictions_m4)) # convert to numeric to calculate ROC

# Make predictions for model 4 on the test set
test_predictions_m4 <- predict(model4, newdata = validation_merge_ds, type = "response")
test_predictions_m4 <- as.numeric(as.character(test_predictions_m4)) # convert to numeric to calculate ROC

# Calculate AUC for model 4 for train set
train_roc_m4 <- roc(train_merge_ds$TARGET, train_predictions_m4)
train_auc_m4 <- auc(train_roc_m4)
print(paste("AUC for train set:", train_auc_m4)) # AUC = 0.54

# Calculate AUC for model 4 for test set
test_roc_m4 <- roc(validation_merge_ds$TARGET, test_predictions_m4)
test_auc_m4 <- auc(test_roc_m4)
print(paste("AUC for test set:", test_auc_m4)) # AUC = 0.54

# Train Random Forest model 5
model5 <- randomForest(TARGET ~ EXT_SOURCE_3 + EXT_SOURCE_2 + DAYS_BIRTH + DAYS_CREDIT_UPDATE + DAYS_CREDIT + DAYS_ENDDATE_FACT + AMT_CREDIT + AMT_GOODS_PRICE + AMT_ANNUITY.x + AMT_INCOME_TOTAL + DAYS_CREDIT_ENDDATE + DAYS_ID_PUBLISH + CODE_GENDER + DAYS_EMPLOYED + CREDIT_TYPE + OWN_CAR_BIN + NAME_INCOME_TYPE + AMT_CREDIT_SUM + STATUS + EXT_SOURCE_1, data = train_merge_ds,
                   	ntree = 500,
                   	mtry = sqrt(ncol(train_merge_ds) - 1),
                   	importance = TRUE)

# Evaluate model 5
# Make predictions for model 5 on the train set
train_predictions_m5 <- predict(model5, newdata = train_merge_ds, type = "response")
train_predictions_m5 <- as.numeric(as.character(train_predictions_m5)) # convert to numeric to calculate ROC

# Make predictions for model 5 on the test set
test_predictions_m5 <- predict(model5, newdata = validation_merge_ds, type = "response")
test_predictions_m5 <- as.numeric(as.character(test_predictions_m5)) # convert to numeric to calculate ROC

# Calculate AUC for model 5 for train set
train_roc_m5 <- roc(train_merge_ds$TARGET, train_predictions_m5)
train_auc_m5 <- auc(train_roc_m5)
print(paste("AUC for train set:", train_auc_m5)) # AUC = 1.00

# Calculate AUC for model 5 for test set
test_roc_m5 <- roc(validation_merge_ds$TARGET, test_predictions_m5)
test_auc_m5 <- auc(test_roc_m5)
print(paste("AUC for test set:", test_auc_m5)) # AUC = 0.61

# Train Random Forest model 6
model6 <- randomForest(TARGET ~ EXT_SOURCE_3 + EXT_SOURCE_2 + DAYS_BIRTH + DAYS_CREDIT_UPDATE + DAYS_CREDIT + DAYS_ENDDATE_FACT + AMT_CREDIT + AMT_GOODS_PRICE + AMT_ANNUITY.x + AMT_INCOME_TOTAL + DAYS_CREDIT_ENDDATE + DAYS_ID_PUBLISH + CODE_GENDER + DAYS_EMPLOYED + CREDIT_TYPE + OWN_CAR_BIN + NAME_INCOME_TYPE + AMT_CREDIT_SUM + STATUS + EXT_SOURCE_1, data = train_merge_ds,
                   	ntree = 500,
                   	importance = TRUE)

# Evaluate model 6
# Make predictions for model 6 on the train set
train_predictions_m6 <- predict(model6, newdata = train_merge_ds, type = "response")
train_predictions_m6 <- as.numeric(as.character(train_predictions_m6)) # convert to numeric to calculate ROC

# Make predictions for model 6 on the test set
test_predictions_m6 <- predict(model6, newdata = validation_merge_ds, type = "response")
test_predictions_m6 <- as.numeric(as.character(test_predictions_m6)) # convert to numeric to calculate ROC

# Calculate AUC for model 6 for train set
train_roc_m6 <- roc(train_merge_ds$TARGET, train_predictions_m6)
train_auc_m6 <- auc(train_roc_m6)
print(paste("AUC for train set:", train_auc_m6)) # AUC = 1.00

# Calculate AUC for model 6 for test set
test_roc_m6 <- roc(validation_merge_ds$TARGET, test_predictions_m6)
test_auc_m6 <- auc(test_roc_m6)
print(paste("AUC for test set:", test_auc_m6)) # AUC = 0.61
```

